Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-10-24 pkg/disttree/inst/draft/draft.Rnw
added supplements as references
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2363 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 21db45aee2f2a8bfcf6b109c313cb0ef96f1d2f9

# add labels
text(x = -1.7, y = lh1, labels = "2009", col = pal[1], cex = 0.8)     # -0.8
text(x = -1.7, y = lh2, labels = "2010", col = pal[3], cex = 0.8)     # -0.8
text(x = -1.7, y = lh3, labels = "2011", col = pal[2], cex = 0.8)     # -0.8
text(x = -1.7, y = lh4, labels = "2012", col = pal[4], cex = 0.8)     # -0.8
@
\caption{\label{fig:axams-dist}\hlc{Total precipitation predictions by a 
distributional forest at station Axams for July~24 in 2009, 2010, 2011 and 
2012 learned on data from 1985--2008. Observations are non-negative and 
modeled by a Gaussian distribution left-censored at zero.
The observations are depicted by crosses and the predicted point mass
from the model by filled circles.}}
\end{figure}

@@

\item \emph{EMOS:} \hcl{The basic} ensemble model output statistics
  models use the ensemble mean of total precipitation as regressor in the location
  submodel and the corresponding ensemble standard deviation in the scale submodel.
  The parameters are estimated by maximum likelihood, using an identity
  link for the location part and a log link for the scale part
  \citep[following the advice of][]{Gebetsberger+Messner+Mayr:2017}.

@@

To assess calibration Figure~\ref{fig:axams-qq} shows residual QQ~plots for
out-of-sample predictions (2009--2012) from the different models trained on
1985--2008. Due to the point masses at zero 100 draws from the randomized
quantile residuals \citep{Dunn+Smyth:1996} are plotted in semi-transparent
gray. Overall, the randomized quantile residuals conform quite well with the
theoretical standard normal quantile (i.e., form a straight line close to the 
diagonal), indicating that all four models are sufficiently well calibrated. 
\hlc{This is also supported by the corresponding probability integral transform}
(PIT, \citealt{Gneiting+Balabdaoui+Raftery:2007}) \hlc{histograms for station
Axams in Supplement 2 (Stationwise Evaluation) which contains a more 
detailed explanation of residual QQ plots and PIT histograms.}

@@

\subsubsection{Distributional specifications}
%% choice of distribution mit Verweis auf Supplement distributions
\hlc{Choosing an adequate distributional family is a crucial step and forms
the basis required to build a well fitting model. Therefore, this
decision should be made thoughtfully. 
In the presented application we applied a Gaussian distribution left-censored 
at zero as this has proven to be a well working choice in
earlier literature} \citep{Stauffer+Mayr+Messner:2017}. 
\hlc{To test for robustness against distributional misspecifation 
different distributional assumptions have been considered as well. 
In Supplement 1 (Different Response Distributions) the exact
same evaluations as in} Section~\ref{sec:all} \hlc{are carried out for 15
observation stations, once employing a Gaussian hurdle model and once
with a logistic distribution left-censored at zero in order to account
for heavier tails. Under both assumptions the obtained results 
are similar to those presented in the paper applying a Gaussian distribution
left-censored at zero. In some settings the hurdle model leads to slight 
improvements, but this comes with more variability (especially for EMOS and the 
boosted GAMLSS). For the boosted GAMLSS the left-censored logistic distribution
shows slightly better results regarding CRPS skill score, however, only within 
a range of less than $2\%$ and not for all stations. 
Altogether, the distributional forest has proven to be 
very robust against distributional misspecifications.}

@@

\cite{Hastie+Tibshirani+Friedman:2001} and \cite{Breiman:2001} \hlc{recommend to 
build full-grown trees. Therefore, low values should be chosen for the stopping 
criteria} \code{minsplit} and \code{minbucket} \hcl{together with a high significance 
level} \code{alpha}.
\hlc{However, in a distributional model it also has to be considered that a 
minimal number of observations is required to allow for a reasonable 
estimation of parameters in each node. For that reason, the minimal number
of observations for which statistical tests are performed in the 
distributional forest in} Section~\ref{sec:precipitation} is set to
50 (\code{minbucket}) \hlc{and the minimal number of observations on which
parameters are estimated is set to 20} (\code{minsplit}).

@@

\subsubsection*{East Tyrol}
Figure~\ref{fig:all-map} \hlc{shows that, contrary to the rest of Tyrol, 
for the majority of stations in East Tyrol the prespecified GAMLSS performs
best in this particular setting (learning data: 1985--2008, testing data:
2009--2012). East Tyrol is located south of the main Alpine Ridge and 
exhibits different climatological characteristics than the part north of 
the Alpine Ridge. It seems that the choice of covariates and their interactions 
we made for the prespecified GAMLSS method is particularly adequate for the 
region of East Tyrol for this specific analysis. However, the advantages of 
the prespecified GAMLSS models in this particular setting (training: 1985--2008, 
validation: 2009--2012) even out in a cross-validation framework for most of 
the stations in East Tyrol. These results are presented in Supplement 2 
(Stationwise Evaluation) where the exact same evaluations as in} 
Section~\ref{sec:axams} \hlc{for station Axams have been carried out 
for 14 other stations.}

\subsubsection*{Choice of station Axams}
%% choice of station Axams mit Verweis auf Supplement stationwise
\hlc{Axams was chosen for the more extensive evaluations presented
in this paper as it is the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked, whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is also confirmed by 
the extended analysis in Supplement 2 (Stationwise Evaluation)
where the 15 considered stations also include some of those where 
the forest does not perform better than the other models in the setting 
presented in} Figure~\ref{fig:all-map}. \hlc{It shows that the differences in 
performance even out in the cross-validation framework where the forest 
performs at least equally well compared to the best of the other models 
at most stations.}

@@

The fitted distributional forest for July 24 and
observation station Axams (including Figure~\ref{fig:axams-dist})
is reproducible using 
\code{demo("RainAxams", package = "disttree")}.
This also includes fitting the other zero-censored
Gaussian models considered in this paper and generating the corresponding 
QQ~plots (Figure~\ref{fig:axams-qq}) and PIT histograms 
(Supplement 2: Stationwise Evaluation).
Full replication of all results can be obtained with
\code{demo("RainTyrol", package = "disttree")}
requiring the companion \textsf{R} package \textbf{RainTyrol} (version~\Sexpr{packageVersion("RainTyrol")}),
also available within the \textsf{R}-Forge~project.
The results presented in Supplement 1 (Different Response Distributions) and 
Supplement 2 (Stationwise Evaluation) can be reproduced using
\code{demo("RainDistributions", package = "disttree")} and 
\code{demo("RainStationwise", package = "disttree")} respectively.

>>>>

# add labels
text(x = -1.7, y = lh1, labels = "2009", col = pal[1], cex = 0.8)     # -0.8
text(x = -1.7, y = lh2, labels = "2010", col = pal[3], cex = 0.8)     # -0.8
text(x = -1.7, y = lh3, labels = "2011", col = pal[2], cex = 0.8)     # -0.8
text(x = -1.7, y = lh4, labels = "2012", col = pal[4], cex = 0.8)     # -0.8
@
\caption{\label{fig:axams-dist}\protect\hlc{Total precipitation predictions by a 
distributional forest at station Axams for July~24 in 2009, 2010, 2011 and 
2012 learned on data from 1985--2008. Observations are non-negative and 
modeled by a Gaussian distribution left-censored at zero.
The observations are depicted by crosses and the predicted point mass
from the model by filled circles.}}
\end{figure}

@@

\item \emph{EMOS:} \hlc{The basic} ensemble model output statistics
  models use the ensemble mean of total precipitation as regressor in the location
  submodel and the corresponding ensemble standard deviation in the scale submodel.
  The parameters are estimated by maximum likelihood, using an identity
  link for the location part and a log link for the scale part
  \citep[following the advice of][]{Gebetsberger+Messner+Mayr:2017}.

@@

To assess calibration Figure~\ref{fig:axams-qq} shows residual QQ~plots for
out-of-sample predictions (2009--2012) from the different models trained on
1985--2008. Due to the point masses at zero 100 draws from the randomized
quantile residuals \citep{Dunn+Smyth:1996} are plotted in semi-transparent
gray. Overall, the randomized quantile residuals conform quite well with the
theoretical standard normal quantile (i.e., form a straight line close to the 
diagonal), indicating that all four models are sufficiently well calibrated. 
\hlc{This is also supported by the corresponding probability integral transform}
(PIT, \citealt{Gneiting+Balabdaoui+Raftery:2007}) \hlc{histograms for station
Axams in Supplement 2} \citep{Schlosser+Hothorn+Stauffer:2018b} \hlc{which 
contains a more detailed explanation of residual QQ plots and PIT histograms.}

@@

\subsubsection{Distributional specifications}
\hlc{Choosing an adequate distributional family is a crucial step and forms
the basis required to build a well fitting model. Therefore, this
decision should be made thoughtfully. 
In the presented application we applied a Gaussian distribution left-censored 
at zero as this has proven to be a well working choice in
earlier literature} \citep{Stauffer+Mayr+Messner:2017}. 
\hlc{To test for robustness against distributional misspecifation 
different distributional assumptions have been considered as well. 
In Supplement 1} \citep{Schlosser+Hothorn+Stauffer:2018a} \hlc{the exact
same evaluations as in} Section~\ref{sec:all} \hlc{are carried out for 15
observation stations, once employing a Gaussian hurdle model and once
with a logistic distribution left-censored at zero in order to account
for heavier tails. Under both assumptions the obtained results 
are similar to those presented in the paper applying a Gaussian distribution
left-censored at zero. In some settings the hurdle model leads to slight 
improvements, but this comes with more variability (especially for EMOS and the 
boosted GAMLSS). For the boosted GAMLSS the left-censored logistic distribution
shows slightly better results regarding CRPS skill score, however, only within 
a range of less than $2\%$ and not for all stations. 
Altogether, the distributional forest has proven to be 
very robust against distributional misspecifications.}

@@

\cite{Hastie+Tibshirani+Friedman:2001} and \cite{Breiman:2001} \hlc{recommend to 
build full-grown trees. Therefore, low values should be chosen for the stopping 
criteria} \code{minsplit} and \code{minbucket} \hlc{together with a high significance 
level} \code{alpha}.
\hlc{However, in a distributional model it also has to be considered that a 
minimal number of observations is required to allow for a reasonable 
estimation of parameters in each node. For that reason, the minimal number
of observations for which statistical tests are performed in the 
distributional forest in} Section~\ref{sec:precipitation} is set to
50 (\code{minbucket}) \hlc{and the minimal number of observations on which
parameters are estimated is set to 20} (\code{minsplit}).

@@

\subsubsection*{East Tyrol}
Figure~\ref{fig:all-map} \hlc{shows that, contrary to the rest of Tyrol, 
for the majority of stations in East Tyrol the prespecified GAMLSS performs
best in this particular setting (learning data: 1985--2008, testing data:
2009--2012). East Tyrol is located south of the main Alpine Ridge and 
exhibits different climatological characteristics than the part north of 
the Alpine Ridge. It seems that the choice of covariates and their interactions 
we made for the prespecified GAMLSS method is particularly adequate for the 
region of East Tyrol for this specific analysis. However, the advantages of 
the prespecified GAMLSS models in this particular setting (training: 1985--2008, 
validation: 2009--2012) even out in a cross-validation framework for most of 
the stations in East Tyrol. These results are presented in Supplement 2} 
\citep{Schlosser+Hothorn+Stauffer:2018b} \hlc{where the exact same evaluations as in} 
Section~\ref{sec:axams} \hlc{for station Axams have been carried out 
for 14 other stations.}

\subsubsection*{Choice of station Axams}
%% choice of station Axams mit Verweis auf Supplement stationwise
\hlc{Axams was chosen for the more extensive evaluations presented
in this paper as it is the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked, whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is also confirmed by 
the extended analysis in Supplement 2} \citep{Schlosser+Hothorn+Stauffer:2018b}
\hlc{where the 15 considered stations also include some of those where 
the forest does not perform better than the other models in the setting 
presented in} Figure~\ref{fig:all-map}. \hlc{It shows that the differences in 
performance even out in the cross-validation framework where the forest 
performs at least equally well compared to the best of the other models 
at most stations.}

@@

The fitted distributional forest for July 24 and
observation station Axams (including Figure~\ref{fig:axams-dist})
is reproducible using 
\code{demo("RainAxams", package = "disttree")}.
This also includes fitting the other zero-censored
Gaussian models considered in this paper and generating the corresponding 
QQ~plots (Figure~\ref{fig:axams-qq}) and PIT histograms 
\citep[Supplement 2]{Schlosser+Hothorn+Stauffer:2018b}.
Full replication of all results can be obtained with
\code{demo("RainTyrol", package = "disttree")}
requiring the companion \textsf{R} package \textbf{RainTyrol} (version~\Sexpr{packageVersion("RainTyrol")}),
also available within the \textsf{R}-Forge~project.
\hlc{The results presented in Supplement 1} \citep{Schlosser+Hothorn+Stauffer:2018a} and 
Supplement 2 \citep{Schlosser+Hothorn+Stauffer:2018b} \hlc{can be reproduced using}
\code{demo("RainDistributions", package = "disttree")} and 
\code{demo("RainStationwise", package = "disttree")} respectively.
